Clustering in a hyperbolic model of complex networks.

### 2019-05-28 introduction.tex
included all comments from Nikolaos and made final changes. Ready for Tobias.

%
%\subsection{Hyperbolic random graphs}\label{ssec:hyperbolic_model}

@@

As observed by Krioukov et al.~\cite{krioukov2010hyperbolic}, and proved rigorously by Gugelmann et al.~\cite{gugelmann2012random}, the degree sequence of the KPKVB model follows a power-law with exponent $2\alpha+1$.
% More specifically, if $N_k$ denotes the number of vertices of degree exactly $k$ then Gugelmann et al.~\cite{gugelmann2012random} have shown that
% 
% \begin{equation}\label{eq:gugeldegseq} 
% N_k = (1+o(1)) \cdot n \cdot p_k \quad \text{ a.a.s., }
% \end{equation}
% 
% \noindent
% where $p_k := 2\alpha \cdot \xi^{2\alpha} \cdot \Gamma^+\left(k - 2\alpha, \xi\right) / k!$
% with, here and in the rest of the paper, $\xi := \frac{4\alpha\nu}{\pi(2\alpha-1)}$ and 
% $\Gamma^+(a,b) := \int_b^\infty t^{a-1} e^{-t} \text{d}t$ the 
% {\em upper incomplete gamma function} and, for $(E_n)_n$ a sequence of events, 
% $E_n$ {\em asymptotically almost surely} (a.a.s.) means that $\Pee(E_n)\to 1$.
%%In particular $X_n = (1+o(1)) Y_n$ is equivalent to $X_n/Y_n \plim 1$.
%In fact~\eqref{eq:gugeldegseq} was shown in~\cite{gugelmann2012random} to hold not only when $k$ is constant but 
%also if $k$ depends on $n$ as long as $k = o( n^\delta )$ for a suitable small constant $\delta = \delta( \alpha )$.
%Standard asymptotic manipulations show that 
%
%$$p_k \sim 2\alpha\xi^{2\alpha} k^{-(2\alpha+1)}, $$
%
%as $k\to\infty$.
%\TM{ \RD{We need to add something like the following} : 
%In Appendix~\ref{} we settle the minor technical point of showing that~\eqref{eq:gugeldegseq} in fact extends 
%to all $k$ with $k \ll n^{1/(2\alpha+1)}$, while 
%$N_k = 0$ a.a.s.~whenever $k \gg n^{1/(2\alpha+1)}$. }
Gugelmann et al.~\cite{gugelmann2012random} also showed that the average degree converges in probability to the constant $8\nu\alpha^2/ \pi (2\alpha-1)^2$, and they showed that the (local) clustering coefficient is non-vanishing in the sense that it is bounded below by a positive constant a.a.s. Here, and in the rest of the paper, for a sequence $(E_n)_n$ of events, $E_n$ {\em asymptotically almost surely} (a.a.s.) means that $\Pee(E_n)\to 1$ as $n \to \infty$.
%Their work however left open the question of whether the clustering coefficient actually converges in probability.

@@

A general expression of the clustering function for KPKVB random graphs is given in~\cite[Equation (59)]{krioukov2010hyperbolic}. The authors conjecture that as $k$ tends to infinity, the clustering function decays as $k^{-1}$. They based this prediction on observations (Figure 8 in \cite{krioukov2010hyperbolic}) in experiments on the infrastructure of the Internet obtained in~\cite{claffy2009internet}. Despite these interesting observations and the attention the KPKVB model has generated since then, the behaviour of the clustering function in KPKVB random graphs had not been completely determined. In particular it has not been established whether it converges as $n\to\infty$ to some suitable limit function, nor how $c(k;G)$ scales with $k$. Theorems~\ref{thm:mainkfixed},~\ref{thm:mainktoinfty} and Proposition~\ref{prop:asymp} below settle these questions. Theorem~\ref{thm:mainkfixed} shows that for each fixed $k$, the value $c(k;G(n;\alpha,\nu))$ converges in probability to a constant $\gamma(k)$ that we again give explicitly as a closed form expression involving $\alpha,\nu$ and several classical special functions. Theorem~\ref{thm:mainktoinfty} extends this result to growing sequences satisfying $k \ll n^{1/(2\alpha+1)}$. Proposition~\ref{prop:asymp} clarifies the asymptotic behavior of the limiting function $\gamma(k)$, as $k\to\infty$. This depends on the parameter $\alpha$, and $\gamma(k)$ only scales with $k^{-1}$ when $\alpha > 3/4$, which corresponds to the exponent of the degree distribution exceeding $5/2$. 
So in particular our findings disprove the abovementioned conjecture of Krioukov et al.~\cite{krioukov2010hyperbolic}.

@@

\begin{theorem}\label{thm:mainktoinfty}
Let $\alpha>\frac12, \nu>0$ be fixed and let $k_n$ be a sequence of integers
satisfying $1 \ll k_n \ll n^{1/(2\alpha+1)}$. Then, writing $G_n := G(n;\alpha,\nu)$, we have
\[
	\E[|c(k_n;G_n)-\gamma(k_n)|] = \smallO{\gamma(k_n)},
\]
as $n \to \infty$, where $\gamma(\cdot)$ is as in Theorem~\ref{thm:mainkfixed}. In particular
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c(k_n; G_n)}{\gamma(k_n)}- 1 \right|} = 0.
\]
\end{theorem}

%We remark that the number of vertices of degree exactly $k$ satisfies $N_k = 0$ a.a.s.~when $k \gg n^{1/(2\alpha+1)}$, see~\TM{ add reference to correct thm / lemma here! }.
%So the range in the above result is essentially to best possible.

@@

Note that Theorem~\ref{thm:mainktoinfty} implies that the clustering function of the KPKVB model scales as $\gamma(k)$ as $k$ grows, whose scaling is given in the above result. In particular, this contradicts the scaling conjectured in~\cite{krioukov2010hyperbolic} for $\alpha \leq \frac34$, and confirms it only for $\alpha > \frac34$.

We remark that simultaneously and independently Stegehuis, van der Hofstad and van Leeuwaarden~\cite{stegehuis2018scale} found a similar, but less detailed result on the $k\to\infty$ scaling of the clustering function in the KPKVB model.

@@

Our results for the local clustering function in particular imply uniform convergence of $c(k; G_n)$ for all $2 \le k \le a_n$ where $a_n \ll n^{\frac{1}{2\alpha + 1}}$. To see this let
\[
	b_n = \arg \max_{2 \le k \le a_n} \Exp{\left|\frac{c(k_n; G_n)}{\gamma(k)} - 1\right|}.
\]
Then $b_n \le a_n \ll n^{\frac{1}{2\alpha + 1}}$ and therefore by Theorem~\ref{thm:local_clustering_hyperbolic}
\[
	\lim_{n \to \infty} \max_{2 \le k \le a_n} \Exp{\left|\frac{c(k; G_n)}{\gamma(k)} - 1\right|} 
	= \lim_{n \to \infty}\Exp{\left|\frac{c(b_n; G_n)}{\gamma(b_n)} - 1\right|} = 0.
\]

\paragraph{Near maximum scaling for $k_n$.}
Our results for the clustering function in the KPKVB model are valid for any sequence of integers $k_n \to \infty$ such that $k_n \ll n^{\frac{1}{2\alpha + 1}}$. Although one would like to have results for any sequence $k_n \le n$, it turns out that $n^{1/(2\alpha + 1)}$ is the near maximum scaling for which Theorem~\ref{thm:clustering_coefficient_hyperbolic} can be true. To see why this is the case note that by definition of the local clustering function~\eqref{eq:def_clustering_function} we have that $c(k_n; G_n) = 0$ if $N_{n}(k_n) = 0$, where $N_n(k_n)$ denote the number of vertices in a KPKVB graph $G(n; \alpha, \nu)$ with degree $k_n$. It follows by Markov's inequality that for any positive function $f : \R_+ \to \R_+$
\begin{align*}
	\Exp{\left|\frac{c(k_n; G_n)}{f(k_n)} - 1\right|} 
	&\ge \Exp{\left|\frac{c(k_n; G_n)}{f(k_n)} - 1\right|\ind{N_n(k_n) = 0}}\\ 
	&\ge \Prob{N_{n}(k_n) = 0} \ge 1 - \Exp{N_n(k_n)}.
\end{align*}
We shall later establish (see Lemma~\ref{lem:diff_Nk_hyperbolic_binomial_poisson}) that $\Exp{N_n(k_n)}$ scales as $n k_n^{-(2\alpha + 1)}$. Therefore if $k_n$ is such that $k_n^{-(2\alpha + 1)} n$ tends to zero as $n \to \infty$ we have that 
\[
	\lim_{n \to \infty} \Exp{N_{n}(k_n)} = 0
\]
and hence
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c(k_n; G_n)}{f(k_n)} - 1\right|} 
	\ge 1 - \lim_{n \to \infty} \Exp{N_{n}(k_n)} = 1 \ne 0,
\]
for any positive function $f$. This implies that we cannot expect a result like that of Theorem~\ref{thm:mainktoinfty} to hold as soon as $k_n \gg n^{\frac{1}{2\alpha + 1}}$.

@@

\PvdH{We should try and add something here. Not necessarily needed for the submission to arXiv though.}

>>>>

@@

As observed by Krioukov et al.~\cite{krioukov2010hyperbolic}, and proved rigorously by Gugelmann et al.~\cite{gugelmann2012random}, the degree sequence of the KPKVB model follows a power-law with exponent $2\alpha+1$.
% More specifically, if $N_k$ denotes the number of vertices of degree exactly $k$ then Gugelmann et al.~\cite{gugelmann2012random} have shown that
% 
% \begin{equation}\label{eq:gugeldegseq} 
% N_k = (1+o(1)) \cdot n \cdot p_k \quad \text{ a.a.s., }
% \end{equation}
% 
% \noindent
% where $p_k := 2\alpha \cdot \xi^{2\alpha} \cdot \Gamma^+\left(k - 2\alpha, \xi\right) / k!$
% with, here and in the rest of the paper, $\xi := \frac{4\alpha\nu}{\pi(2\alpha-1)}$ and 
% $\Gamma^+(a,b) := \int_b^\infty t^{a-1} e^{-t} \text{d}t$ the 
% {\em upper incomplete gamma function} and, for $(E_n)_n$ a sequence of events, 
% $E_n$ {\em asymptotically almost surely} (a.a.s.) means that $\Pee(E_n)\to 1$.
%%In particular $X_n = (1+o(1)) Y_n$ is equivalent to $X_n/Y_n \plim 1$.
%In fact~\eqref{eq:gugeldegseq} was shown in~\cite{gugelmann2012random} to hold not only when $k$ is constant but 
%also if $k$ depends on $n$ as long as $k = o( n^\delta )$ for a suitable small constant $\delta = \delta( \alpha )$.
%Standard asymptotic manipulations show that 
%
%$$p_k \sim 2\alpha\xi^{2\alpha} k^{-(2\alpha+1)}, $$
%
%as $k\to\infty$.
%\TM{ \RD{We need to add something like the following} : 
%In Appendix~\ref{} we settle the minor technical point of showing that~\eqref{eq:gugeldegseq} in fact extends 
%to all $k$ with $k \ll n^{1/(2\alpha+1)}$, while 
%$N_k = 0$ a.a.s.~whenever $k \gg n^{1/(2\alpha+1)}$. }
Gugelmann et al.~\cite{gugelmann2012random} also showed that the average degree converges in probability to the constant $8\nu\alpha^2/ \pi (2\alpha-1)^2$, and they showed that the (local) clustering coefficient is non-vanishing in the sense that it is bounded below by a positive constant a.a.s. Here, and in the rest of the paper, for a sequence $(E_n)_n$ of events, $E_n$ {\em asymptotically almost surely} (a.a.s.) means that $\Prob{E_n} \to 1$ as $n \to \infty$. Here the events $E_n$ might be defined in different probability spaces, in which case $\Pee$ will the corresponding probability measure.
%Their work however left open the question of whether the clustering coefficient actually converges in probability.

@@

A general expression of the clustering function for KPKVB random graphs is given in~\cite[Equation (59)]{krioukov2010hyperbolic}. The authors conjecture that as $k$ tends to infinity, the clustering function decays as $k^{-1}$. They based this prediction on observations (Figure 8 in \cite{krioukov2010hyperbolic}) in experiments on the infrastructure of the Internet obtained in~\cite{claffy2009internet}. Despite these interesting observations and the attention the KPKVB model has generated since then, the behaviour of the clustering function in KPKVB random graphs had not been completely determined. In particular it has not been established whether it converges as $n\to\infty$ to some suitable limit function, nor how $c(k;G(n; \alpha,\nu))$ scales with $k$. Theorems~\ref{thm:mainkfixed},~\ref{thm:mainktoinfty} and Proposition~\ref{prop:asymp} below settle these questions. Theorem~\ref{thm:mainkfixed} shows that for each fixed $k$, the value $c(k;G(n;\alpha,\nu))$ converges in probability to a constant $\gamma(k)$ that we again give explicitly as a closed form expression involving $\alpha,\nu$ and several classical special functions. Theorem~\ref{thm:mainktoinfty} extends this result to growing sequences satisfying $k \ll n^{1/(2\alpha+1)}$. Proposition~\ref{prop:asymp} clarifies the asymptotic behavior of the limiting function $\gamma(k)$, as $k\to\infty$. This depends on the parameter $\alpha$, and $\gamma(k)$ only scales with $k^{-1}$ when $\alpha > 3/4$, which corresponds to the exponent of the degree distribution exceeding $5/2$. 
So in particular our findings disprove the abovementioned conjecture of Krioukov et al.~\cite{krioukov2010hyperbolic}.

@@

\begin{theorem}\label{thm:mainktoinfty}
Let $\alpha>\frac12, \nu>0$ be fixed and let $k_n$ be a sequence of non-negative integers
satisfying $1 \ll k_n \ll n^{1/(2\alpha+1)}$. Then, writing $G_n := G(n;\alpha,\nu)$, we have
\[
	\E[|c(k_n;G_n)-\gamma(k_n)|] = \smallO{\gamma(k_n)},
\]
as $n \to \infty$, where $\gamma(\cdot)$ is as in Theorem~\ref{thm:mainkfixed}. In particular
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c(k_n; G_n)}{\gamma(k_n)}- 1 \right|} = 0.
\]
\end{theorem}

@@

Note that Theorem~\ref{thm:mainktoinfty} implies that the clustering function of the KPKVB model scales as $\gamma(k)$ as $k$ grows, whose scaling is given in the above result. In particular, this contradicts the scaling conjectured in~\cite{krioukov2010hyperbolic} for $\alpha \leq \frac{3}{4}$, and confirms it only for $\alpha > \frac{3}{4}$.

We remark that simultaneously and independently Stegehuis, van der Hofstad and van Leeuwaarden~\cite{stegehuis2019scale} used a completely different technique to obtain a similar, though less detailed, result on the $k\to\infty$ scaling of the clustering function in the KPKVB model.

@@

Our results for the local clustering function in particular imply uniform convergence of $c(k; G_n)$ for all $2 \le k \le a_n$ where $a_n \ll n^{\frac{1}{2\alpha + 1}}$. To see this let
\[
	b_n = \arg \max_{2 \le k \le a_n} \Exp{\left|\frac{c(k; G_n)}{\gamma(k)} - 1\right|}.
\]
Then $b_n \le a_n \ll n^{\frac{1}{2\alpha + 1}}$ and therefore by Theorem~\ref{thm:mainktoinfty}
\[
	\lim_{n \to \infty} \max_{2 \le k \le a_n} \Exp{\left|\frac{c(k; G_n)}{\gamma(k)} - 1\right|} 
	= \lim_{n \to \infty}\Exp{\left|\frac{c(b_n; G_n)}{\gamma(b_n)} - 1\right|} = 0.
\]

\paragraph{Near maximum scaling for $k_n$.}
Our results for the clustering function in the KPKVB model are valid for any sequence of non-negative integers $k_n \to \infty$ such that $k_n \ll n^{\frac{1}{2\alpha + 1}}$. Although one would like to have results for any sequence $k_n \le n-1$, it turns out that $n^{1/(2\alpha + 1)}$ is the near maximum scaling for which Theorem~\ref{thm:clustering_coefficient_hyperbolic} can be true. To see why this is the case note that by definition of the local clustering function~\eqref{eq:def_clustering_function} we have that $c(k_n; G_n) = 0$ if $N_{n}(k_n) = 0$, where $N_n(k_n)$ denotes the number of vertices in a KPKVB graph $G(n; \alpha, \nu)$ with degree $k_n$. It follows by Markov's inequality that for any positive function $f : \R_+ \to \R_+$
\begin{align*}
	\Exp{\left|\frac{c(k_n; G_n)}{f(k_n)} - 1\right|} 
	&\ge \Exp{\left|\frac{c(k_n; G_n)}{f(k_n)} - 1\right|\ind{N_n(k_n) = 0}}\\ 
	&\ge \Prob{N_{n}(k_n) = 0} \ge 1 - \Exp{N_n(k_n)}.
\end{align*}
We shall later establish (see Lemma~\ref{lem:diff_Nk_hyperbolic_binomial_poisson}) that $\Exp{N_n(k_n)}$ scales as $n k_n^{-(2\alpha + 1)}$. Therefore if $k_n$ is such that $k_n^{-(2\alpha + 1)} n$ tends to zero as $n \to \infty$ we have that 
\[
	\lim_{n \to \infty} \Exp{N_{n}(k_n)} = 0
\]
and hence
\[
	\lim_{n \to \infty} \Exp{\left|\frac{c(k_n; G_n)}{f(k_n)} - 1\right|} 
	\ge 1 - \lim_{n \to \infty} \Exp{N_{n}(k_n)} = 1 \ne 0,
\]
for any positive function $f$. This implies that we cannot expect a result like that of Theorem~\ref{thm:mainktoinfty} to hold as soon as $k_n \gg n^{\frac{1}{2\alpha + 1}}$.

@@

It follows from Proposition~\ref{prop:asymp} that there is a transition in the scaling of the local clustering function at $\alpha = 3/4$. This corresponds to an exponent $5/2$ for the probability mass function of the degree distribution. This transition is different from those often observed for networks with scale-free degree distributions, where transitions occur at integer values of the exponent. At this point, it is unclear what the underlying reason is for the appearance of the transition at this half exponent. Interestingly, a similar transition point has also been observed for both majority vote models~\cite{chen2015critical} and flocking dynamics~\cite{miguel2018effects} on networks with scale-free degree degree distributions.

### 2019-05-28 paper.tex
processed most of Nikolaos comments on sections 5-7
included all comments from Nikolaos and made final changes. Ready for Tobias.

% Redefines the space between sections in the TOC so that the entire thing fits on one page. 
% Comment this out when not using the tocloft package
\setlength{\cftbeforesecskip}{5pt}

@@

\newcommand{\hsym}[1]{h_{\H \bigtriangleup \Pcal}(#1)}
\newcommand{\EdgeDiff}{E_{\H \setminus \mathcal{P}}}
\newcommand{\expH}{\overline{n}_{\Po}(k_n)}
\newcommand{\expP}{\overline{n}_{\infty}(k_n)}
\newcommand{\Pop}[2]{\mathrm{Po} (#2;#1)}
\newcommand{\Nsymdiff}{N_{\H \bigtriangleup \Pcal}(k_n)}
\renewcommand{\deg}{\mathrm{deg}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Don't mess with the tobimacros !!!!!							  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\eR}[0]{\ensuremath{{\mathbb R}}}
\newcommand{\eN}[0]{\ensuremath{{\mathbb N}}}
\newcommand{\Ee}[0]{\ensuremath{{\mathbb E}}}
\newcommand{\Haa}{\ensuremath{{\mathbb H}}}
\newcommand{\GPo}[0]{\ensuremath{G_{\text{Po}}}}
\newcommand{\Gbox}[0]{\ensuremath{G_{\text{box}}}}
\newcommand{\GPoH}{\ensuremath{G_{Po,H}}}
\newcommand{\GboxH}{\ensuremath{G_{box,H}}}
\newcommand{\Ginf}[0]{\ensuremath{G_{\infty}}}
\newcommand{\isd}{\stackrel{\text{d}}{=}}
\newcommand{\Xcal}{\ensuremath{{\mathcal X}}}
\newcommand{\Ycal}{\ensuremath{{\mathcal Y}}}
\newcommand{\Qcal}{\ensuremath{{\mathcal Q}}}
\newenvironment{proofof}[1]{\vspace{1ex}\noindent{\bf Proof of #1:}}{\hspace*{\fill}$\blacksquare$\vspace{1ex}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%								For leaving comments to each other:								  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Here we show that the clustering coefficient tends in probability to a constant $\gamma$ that we give explicitly as a closed form expression in terms of $\alpha, \nu$ and certain special functions. This improves over earlier work by Gugelmann et al., who proved that the clustering coefficient remains bounded away from zero with high probability, but left open the issue of convergence to a limiting constant. Similarly, we are able to show that $c(k)$, the average clustering coefficient over all vertices of degree exactly $k$, tends in probability to a limit $\gamma(k)$ given explicitly as a closed form expression in terms of $\alpha, \nu$ and certain special functions. We are able to extend this last result also to sequences $(k_n)_n$ where $k_n$ grows as a function of $n$. Our results show that $\gamma(k)$ scales differently, as $k$ grows, for different ranges of $\alpha$. More precisely, $\gamma(k) = \Theta( k^{2 - 4\alpha})$ if $\frac{1}{2} < \alpha < \frac{3}{4}$, $\gamma(k) = \Theta( \log(k) / k )$ if $\alpha=\frac{3}{4}$ and $\gamma = \bigT{k^{-1}}$ when $\alpha > \frac{3}{4}$. These results contradict a prediction of Krioukov et al., which stated that the limiting values $\gamma(k)$ should always scale with $k^{-1}$ as we let $k$ grow. 
\end{abstract}

@@

\paragraph{Acknowledgments} We are grateful to Dmitri Krioukov for pointing us to the problem of local clustering and helpful discussions.

>>>>

% Redefines the space between sections in the TOC so that the entire thing fits on one page. 
% Comment this out when not using the tocloft package
\setlength{\cftbeforesecskip}{7pt}

@@

\newcommand{\hsym}[1]{h_{\H \bigtriangleup \Pcal}(#1)}
\newcommand{\EdgeDiff}{E_{\H \setminus \mathcal{P}}}
\newcommand{\expH}{\overline{n}_{\Po}(k_n)}
\newcommand{\expP}{\overline{n}_{\mathrm{box}}(k_n)}
\newcommand{\Pop}[2]{\mathrm{Po} (#2;#1)}
\newcommand{\Nsymdiff}{N_{\H \bigtriangleup \Pcal}(k_n)}
\renewcommand{\deg}{\mathrm{deg}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Don't mess with the tobimacros !!!!!							  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\eR}[0]{\ensuremath{{\mathbb R}}}
\newcommand{\eN}[0]{\ensuremath{{\mathbb N}}}
\newcommand{\Ee}[0]{\ensuremath{{\mathbb E}}}
\newcommand{\Haa}{\ensuremath{{\mathbb H}}}
\newcommand{\GPo}[0]{\ensuremath{G_{\Po}}}
\newcommand{\Gbox}[0]{\ensuremath{G_{\mathrm{box}}}}
\newcommand{\GPoH}{\ensuremath{G_{Po,H}}}
\newcommand{\GboxH}{\ensuremath{G_{box,H}}}
\newcommand{\Ginf}[0]{\ensuremath{G_{\infty}}}
\newcommand{\isd}{\stackrel{\text{d}}{=}}
\newcommand{\Xcal}{\ensuremath{{\mathcal X}}}
\newcommand{\Ycal}{\ensuremath{{\mathcal Y}}}
\newcommand{\Qcal}{\ensuremath{{\mathcal Q}}}
\newenvironment{proofof}[1]{\vspace{1ex}\noindent{\bf Proof of #1:}}{\hspace*{\fill}$\blacksquare$\vspace{1ex}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%								For leaving comments to each other:								  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Here we show that the clustering coefficient tends in probability to a constant $\gamma$ that we give explicitly as a closed form expression in terms of $\alpha, \nu$ and certain special functions. This improves over earlier work by Gugelmann et al., who proved that the clustering coefficient remains bounded away from zero with high probability, but left open the issue of convergence to a limiting constant. Similarly, we are able to show that $c(k)$, the average clustering coefficient over all vertices of degree exactly $k$, tends in probability to a limit $\gamma(k)$ which we give explicitly as a closed form expression in terms of $\alpha, \nu$ and certain special functions. We are able to extend this last result also to sequences $(k_n)_n$ where $k_n$ grows as a function of $n$. Our results show that $\gamma(k)$ scales differently, as $k$ grows, for different ranges of $\alpha$. More precisely, there exists constants $Q_{\alpha,\nu}$ depending on $\alpha$ and $\nu$, such that as $k \to \infty$, $\gamma(k) \sim Q_{\alpha,\nu} k^{2 - 4\alpha}$ if $\frac{1}{2} < \alpha < \frac{3}{4}$, $\gamma(k) \sim Q_{\alpha,\nu} \log(k) k^{-1} $ if $\alpha=\frac{3}{4}$ and $\gamma = Q_{\alpha,\nu} k^{-1}$ when $\alpha > \frac{3}{4}$. These results contradict a prediction of Krioukov et al., which stated that the limiting values $\gamma(k)$ should always scale with $k^{-1}$ as we let $k$ grow. 
\end{abstract}

@@

\paragraph{Acknowledgments} We are grateful to Dmitri Krioukov for pointing us to the problem of local clustering in this model and his helpful insights during discussions of the topic.

### 2019-05-28 proof_overview_finite_k.tex
included all comments from Nikolaos and made final changes. Ready for Tobias.

Imagine that we have an infinite supply of i.i.d.~points $u_1, u_2, \dots$ in the hyperbolic plane $\Haa$ chosen according to the $(\alpha, R)$-quasi uniform distribution. In the standard KPKVB random graph $G(n;\alpha,\nu)$ we take $u_1,\dots, u_n$ as our vertex set and add edges between points at hyperbolic distance at most $R = 2\log(n/\nu)$. In the {\em Poissonized} KPKVB random graph $\GPo := \GPo(n;\alpha,\nu)$, we instead take $N\isd \Po(n)$, a Poisson random variable 
with mean $n$, independent of our i.i.d.~sequence of points and let the vertex set be $u_1,\dots, u_N$ and add edges 
according to the same rule as before. Equivalently, we could say that the vertex set consists of the points of a Poisson point process with intensity function $n g$, where $g$ denotes the probability density of the $(\alpha,R)$-quasi uniform distribution. That is,
\begin{equation}\label{eq:def_quasi_uniform_density}
	g(r,\theta) = \frac{\alpha\sinh(\alpha r)}{2\pi(\cosh(\alpha R) - 1)} \cdot 1_{\{0\leq r\leq R, -\pi<\theta\leq \pi\}}.
\end{equation}

@@

In the remainder of this paper we will write $\BallHyp{p}$ to denote the image under $\Psi$ of the ball of hyperbolic radius $R$ around the point 
$\Psi^{-1}(p)$ for $p \in \Rcal$, i.e. 
\[
	\BallHyp{p} := 
	%\left\{p^\prime := \Psi(u) \, : \, u \in \Dcal_{R_n} \text{ and } d_\H(\Psi^{-1}(p),u) \le R_n\right\}.
	\Psi\left[ \left\{ u \in \Haa : 
	d_\H(\Psi^{-1}(p),u), d_\H(O,u) \le R \right\}\right] \subset \Rcal.
\]

@@

>>>>

Imagine that we have an infinite supply of i.i.d.~points $u_1, u_2, \dots$ in the hyperbolic plane $\Haa$ chosen according to the $(\alpha, R)$-quasi uniform distribution. In the standard KPKVB random graph $G(n;\alpha,\nu)$ we take $u_1,\dots, u_n$ as our vertex set and add edges between points at hyperbolic distance at most $R = 2\log(n/\nu)$. In the \emph{Poissonized} KPKVB random graph $\GPo := \GPo(n;\alpha,\nu)$, we instead take $N\isd \Po(n)$, a Poisson random variable 
with mean $n$, independent of our i.i.d.~sequence of points and let the vertex set be $u_1,\dots, u_N$ and add edges 
according to the same rule as before. Equivalently, we could say that the vertex set consists of the points of a Poisson point process with intensity function $n g$, where $g$ denotes the probability density of the $(\alpha,R)$-quasi uniform distribution. That is,
\begin{equation}\label{eq:def_quasi_uniform_density}
	g(r,\theta) = \frac{\alpha\sinh(\alpha r)}{2\pi(\cosh(\alpha R) - 1)} \cdot 1_{\{0\leq r\leq R, -\pi<\theta\leq \pi\}}.
\end{equation}

@@

In the remainder of this paper we will write $\BallHyp{p}$ to denote the image under $\Psi$ of the ball of hyperbolic radius $R$ around the point 
$\Psi^{-1}(p)$ for $p \in \Rcal$, i.e. 
\[
	\BallHyp{p} := 
	%\left\{p^\prime := \Psi(u) \, : \, u \in \Dcal_{R} \text{ and } d_\H(\Psi^{-1}(p),u) \le R\right\}.
	\Psi\left[ \left\{ u \in \Haa : 
	d_\H(\Psi^{-1}(p),u), d_\H(O,u) \le R \right\}\right] \subset \Rcal.
\]

@@

\begin{remark}[Notion convention for points]
We will often be working with the finite box graph $\GPo$ or the infinite graph $\Ginf$, whose nodes are points in $\R \times \R_+$. For any point $p \in \R \times \R_+$ we will always use $p = (x,y)$. When considering different points $p,p^\prime \in \R \times \R_+$, we will use primed coordinates to refer to $p^\prime$, i.e. $p^\prime = (x^\prime ,y^\prime)$, and similar with subscripts, i.e. $p_i = (x_i,y_i)$.
\end{remark}
